---
title: "Assignment 1"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center', out.width = '70%', fig.asp = .70)
library(tidyverse)
```

# Question 1

1.  Run `ggplot(data = mpg)`. What do you see?

2.  Make a scatterplot of `hwy` vs `cyl`.

3.  What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?

# Question 2

1.  What's gone wrong with this code? Why are the points not blue?

```{r, eval = FALSE}
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```
2.  Which variables in `mpg` are categorical? Which variables are continuous? 
    (Hint: type `?mpg` to read the documentation for the dataset). How can you see this information when you run `mpg`?

3.  Map a continuous variable to `color`, `size`, and `shape`. How do these aesthetics behave differently for categorical vs. continuous variables? 
    
4.  What happens if you map the same variable to multiple aesthetics? 

5.  What does the `stroke` aesthetic do? What shapes does it work with? (Hint: use `?geom_point`)
    
6.  What happens if you map an aesthetic to something other than a variable name, like `aes(colour = displ < 5)`?  
    
# Question 3

1.  What happens if you facet on a continuous variable?

2.  What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean? How do they relate to this plot?
    
```{r, eval = FALSE}
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = drv, y = cyl))
```

3.  What plots does the following code make? What does `.` do?

```{r eval = FALSE}
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) +
      facet_grid(drv ~ .)
    
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) +
      facet_grid(. ~ cyl)
```

4.  Take the first faceted plot in this section:

```{r, eval = FALSE}
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) + 
      facet_wrap(~ class, nrow = 2)
```

What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?
    
5.  Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other options control the layout of the individual panels? Why doesn't `facet_grid()` have `nrow` and `ncol` arguments?

6.  When using `facet_grid()` you should usually put the variable with more unique levels in the columns. Why?

# Question 4

1.  What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

2.  Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.
    
```{r, eval = FALSE}
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
      geom_point() + 
      geom_smooth(se = FALSE)
```

3.  What does `show.legend = FALSE` do?  What happens if you remove it? 

4.  What does the `se` argument to `geom_smooth()` do?


5.  Will these two graphs look different? Why/why not?

```{r, eval = FALSE}
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point() + 
      geom_smooth()
    
    ggplot() + 
      geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

6.  Recreate the R code necessary to generate the following graphs.

```{r echo = FALSE, fig.width = 3, out.width = "50%", fig.align = "default", message = FALSE}
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point() + 
      geom_smooth(se = FALSE)
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_smooth(aes(group = drv), se = FALSE) +
      geom_point()
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
      geom_point() + 
      geom_smooth(se = FALSE)
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point(aes(color = drv)) + 
      geom_smooth(se = FALSE)
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point(aes(color = drv)) +
      geom_smooth(aes(linetype = drv), se = FALSE)
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point(size = 4, colour = "white") + 
      geom_point(aes(colour = drv))
```

# Question 5

1.  What is the default geom associated with `stat_summary()`? How could you rewrite the previous plot (Hadley Wickham, *R for Data Science*, page 25) to use that geom function instead of the stat function?

2.  What does `geom_col()` do? How is it different to `geom_bar()`?

3.  Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?

4.  What variables does `stat_smooth()` compute? What parameters control its behaviour?

# Question 6

1.  What is the problem with this plot? How could you improve it?

```{r}
    ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
      geom_point()
```

2.  What parameters to `geom_jitter()` control the amount of jittering?

3.  Compare and contrast `geom_jitter()` with `geom_count()`.

4.  What's the default position adjustment for `geom_boxplot()`? Create a visualisation of the `mpg` dataset that demonstrates it.
