---
title: "Advanced Statistics HW1"
author: "Lisi(201900101)"
date:  "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,           # show code
  warning = FALSE,       # don't show warnings
  message = FALSE,       # don't show messages (less serious warnings)
  cache = FALSE,         # set to TRUE to save results from last compilation
  fig.align = 'center',  # center figures
  out.width = '75%',     # Set figure size to 70% of textline
  fig.asp = .75          # width-height ratio 
)
  library(tidyverse)     # Load libraries you always use here
  set.seed(2018)         # make random results reproducible
```

## Exercises 1
Let $Y$ be a normal distributed random variable,  $\mu$ is the mean, and $\sigma$ is the standard deviation. What is the probability density function of $Y$?

**Answer:**

Denote $Y \sim N(\mu,\sigma^2)$, then the pdf of $Y$ is 

$$f(y|\mu, \sigma) = \frac{1}{\sqrt{2 \pi \sigma^2}} \text{exp} \Bigg (- \frac{(y - \mu)^2}{2 \sigma^2} \Bigg)$$

## Exercises 2.
Display the dataset `mpg` with `kable`.

**Answer:**

The table of dataset `mpg` with `kable` can be given as below: 

```{r}
knitr::kable(head(mpg),
   caption = 'Table of the first several data', booktabs = TRUE)
```

## Exercises 3
Plot the `hwy` against `displ` in dateset `mpg`.

**Answer:**
 
The code and figure are given as follows
```{r plotmpg, fig.cap="ggplot of hwy vs displ"}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy))
```


